---
title: "CRP City of Decatur Hobo temps_ August 2023"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}


library(ggplot2)
library(tidyverse)
library(lubridate)

```
## Read in data
Stationary HOBO U23 sensor data from interior and exterior/edge of Lullwater Forest and the Outdoor Activity center - late June - early Aug 2021.
```{r echo=FALSE, include = FALSE}



ASCQuad <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/ASC_quad.csv")
head(ASCQuad)
type<-factor("Sun")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
ASCQuad<- data.frame(ASCQuad,type,location)
head(ASCQuad)

##########
DecaturFirst <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Decatur_First.csv")
head(DecaturFirst)
type<-factor("Sun")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
DecaturFirst<- data.frame(DecaturFirst,type,location)
head(DecaturFirst)


##########
Cemetery <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Decatur_First.csv")
head(Cemetery)
type<-factor("Shade")
location<-factor("R-60")

#rename variables and add columns for serial no, location, type (Sun or Shade)
Cemetery<- data.frame(Cemetery,type,location)
head(Cemetery)

MARTA <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Deer_and_Dove.csv")
head(MARTA)
type<-factor("Sun")
location<-factor("C-2")

#rename variables and add columns for serial no, location, type (Sun or Shade)
MARTA<- data.frame(MARTA,type,location)
head(MARTA)


DHSHoward <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/DHS_howard.csv")
head(DHSHoward)
type<-factor("Sun")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
DHSHoward<- data.frame(DHSHoward,type,location)
head(DHSHoward)

Harmony <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Harmony_park.csv")
head(Harmony)
type<-factor("Sun")
location<-factor("C-1")

#rename variables and add columns for serial no, location, type (Sun or Shade)
Harmony<- data.frame(Harmony,type,location)
head(Harmony)

DHSBioswale <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/High_School_BioSwale.csv")
head(DHSBioswale)
type<-factor("Shade")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
DHSBioswale <- data.frame(DHSBioswale ,type,location)
head(DHSBioswale )


OakhurstPark <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Oakhurst_park.csv")
head(OakhurstPark)
type<-factor("Shade")
location<-factor("R-60")

#rename variables and add columns for serial no, location, type (Sun or Shade)
OakhurstPark <- data.frame(OakhurstPark ,type,location)
head(OakhurstPark )


StThomasMoore <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/STM.csv")
head(StThomasMoore)
type<-factor("Sun")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
StThomasMoore <- data.frame(StThomasMoore ,type,location)
head(StThomasMoore )


Talley <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Talley.csv")
head(Talley)
type<-factor("Sun")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
Talley <- data.frame(Talley ,type,location)
head(Talley)

ToyPark <- read.csv("~/Desktop/July_Aug_23_COD_heat_sensors/Toy_Park.csv")
head(ToyPark)
type<-factor("Shade")
location<-factor("I")

#rename variables and add columns for serial no, location, type (Sun or Shade)
ToyPark <- data.frame(ToyPark ,type,location)
head(ToyPark)


#OACForest <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/OAC1Interior_6-29-21_8-19-21.csv")
#head(OACForest)
#type<-factor("Forest")
#location<-factor("OAC")

#rename variables and add columns for serial no, location, type (forest or edge)
#oacF<- data.frame(OACForest,type,location)
#head(oacF)

#OAC Edge dataset with type = edge and add location
#OACEdge <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/OAC2Exterior_7-7-21_8-19-21.csv")
#type<-factor("Edge")
#location<-factor("OAC")
#oacEdge<- data.frame(OACEdge,type,location)


#Lullwater datasets
#LullwaterEdge <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/LullwaterExterior_stonePN4-Emory-4-22-21_0_8-19-21.csv")
#type<-factor("Edge")
#location<-factor("Lullwater")
#lullEdge<- data.frame(LullwaterEdge,type,location)


#LullwaterInterior <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/LullwaterInterior1_stoneOH3_6-29-21_8-19-21.csv")
#type<-factor("Forest")
#location<-factor("Lullwater")
#lullF<- data.frame(LullwaterInterior,type,location)

#Fernbank dataset
#FernbankInterior <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/FernbankSlope_7.10.21_8-19-21.csv")
#type<-factor("Forest")
#location<-factor("Fernbank")
#fernF<- data.frame(FernbankInterior,type,location)

ls(Cemetery)

Cemetery <- Cemetery %>%
  select(X., Date.Time..EST.EDT., Ch.1...Temperature.....C., Ch.2...RH......, Dew.Point.....C., type, location )

DecaturFirst <- DecaturFirst %>%
  select(X., Date.Time..EST.EDT., Ch.1...Temperature.....C., Ch.2...RH......, Dew.Point.....C., type, location )

##Concatenate all 4 datasets by row
alldata <- as.data.frame(rbind(ASCQuad, Cemetery,DecaturFirst,DHSBioswale,DHSHoward,Harmony,MARTA, OakhurstPark, StThomasMoore,Talley, ToyPark))



alldata$Date.Time..EST.EDT. <- mdy_hms(alldata$Date.Time..EST.EDT.)
str(alldata)


#alldata$longdate <- parse_date_time(alldata$Date.Time..EST.EDT., orders = mdy)

#class(alldata$Date.Time..EST.EDT.)
#alldata$longdate <- mdy(alldata$Date.Time..EST.EDT.)

#alldata$longdate<- as_datetime(alldata$Date.Time..EST.EDT., format = "mdy_hms")
#str(alldata)

#https://rpubs.com/cgb-grupstra/moorea-hobo-20190314
#IF need to separate time and date into 2 columns
hobotemplongdate <- tidyr::separate(alldata, 'Date.Time..EST.EDT.',
                                    into = c('longdate', 'time'),
                                   sep= ' ') 

head(hobotemplongdate)



hobotemplongdate$longdate <- as.Date(hobotemplongdate$longdate)



```

## Plotting across dates 
using code from https://rpubs.com/cgb-grupstra/moorea-hobo-20190314
```{r echo=FALSE}

ntempgraph <- ggplot(data=hobotemplongdate, 
    aes(x=longdate, 
    y=Ch.1...Temperature.....C.)) +
    geom_point(size=1, alpha = 1/10)+ theme_bw()+
    facet_grid(cols=vars(location))+ 
    theme(axis.text.x = element_text(angle=65, margin = margin(t=20, r=100)))+
    labs(title="Raw temperature data", y="Temperature (°C)", x="Date")+
    guides(x = guide_axis(angle=65))+
  scale_x_date(date_breaks = '3 months')
ntempgraph


#install.packages('labeling') 
#install.packages('farver')
library(labeling)
library(farver)
```


```{r include=FALSE}

#separating day/month/year 

hobofull <- hobotemplongdate %>%
  tidyr::separate('longdate',
                  into = c('year', 'month', 'day'),
                  sep= '-',
                  remove = FALSE)
head(hobofull)
```


## Daily temps
Calculating mean temp for each date
```{r mean temps}
hobofullmean <- hobofull %>%
  group_by(year,month, day,type,location,longdate)%>%
  summarise(meantemp = mean(Ch.1...Temperature.....C.))

head(hobofullmean)

```

## Plotting mean temps
```{r echo=FALSE}

#hobofullmean$month <- factor(hobofullmean$month, levels=c("06", "07", "08"))
meanplot <- ggplot(hobofullmean, aes(x=longdate, y=meantemp))+
  geom_smooth(aes())+
    geom_point(aes(),alpha = 1/5)+
  facet_grid(cols=vars(location))+
    theme_bw()+
    labs(title= "Daily temperature means", y="Daily mean temperature (°F) with 95% CI", x="Date")+
    guides(x = guide_axis(angle=65))+
  scale_x_date(date_breaks = '3 months')
   
    
meanplot

```

## Calculating daily temp ranges and min/max
```{r temp range}

hobofullrange <- hobofull %>%
  group_by(year, month, day, type, location, longdate)%>%
  summarise(min_temp = min(Ch.1...Temperature.....C.), max_temp = max(Ch.1...Temperature.....C.))%>%
  mutate(range = max_temp-min_temp)

head(hobofullrange, digits=5)

```

## Plotting temp ranges
```{r echo=FALSE}

rangeplot <- ggplot(data=hobofullrange, 
                   aes(x=longdate, 
                       y=range)) +
  geom_point()+ theme_bw()+
  theme(axis.text.x = element_text(margin = margin(t=5)))+
  facet_grid(cols=vars(location))+
  labs(title="Daily temperature range", y="Daily temperature range (°C) with 95% CI ", x="Date")+
  guides(x = guide_axis(angle=65))+
  scale_x_date(date_breaks = '3 months')
rangeplot

```

## Plotting max and min
```{r echo=FALSE}

hobofullrange <- hobofull %>%
  group_by(year, month, day, type, location, longdate)%>%
  summarise(min_temp = min(Ch.1...Temperature.....C.), max_temp = max(Ch.1...Temperature.....C.))%>%
 mutate(range = max_temp-min_temp)

minmaxplot <- ggplot(hobofullrange, aes(x=longdate))+
  geom_point(aes(y=min_temp, color = "blue"))+
    geom_point(aes(y=max_temp, color = "red"))+
  facet_grid(cols=vars(location))+
    theme_bw()+
    labs(title= "Daily temperature Min/Max", y="Daily Max and Min temperature (°F) with 95% CI", x="Date", color = "Min/Max Temps")+
   scale_color_manual(labels = c("Daily Min Temp", "Daily Max Temp"), values = c("blue", "red")) +
    guides(x = guide_axis(angle=65))+
  scale_x_date(date_breaks = '3 months')
minmaxplot
   
#    geom_point(aes(colour=type),alpha = 1/5)+
```

## Boxplots - temps per moth by location and type
median, 25%, 75%, and min and max
```{r echo=FALSE}

#hobofullrangemeans<- hobofullrange %>%
#  group_by(year, month,type, location)%>%
#  summarise(meanrange = mean(range))
#nrow(hobofullrangemeans)

hrmplot <- ggplot(hobofullmean, aes(x=month, y=meantemp,colour=year))+
  geom_boxplot() + theme_bw()+
  facet_grid(cols=vars(location))+
  labs(title= "Monthly average temps ", y="Temperature (°C)", x="Date") +
  guides(x = guide_axis(angle=65))
   

hrmplot

```

```{r}
# Heat Index

ls(hobotemplongdate)

hobotemplongdate$HI_uncorrected <- -42.379 + 2.04901523* hobotemplongdate$Ch.1...Temperature.....C.+ 10.14333127*hobotemplongdate$Ch.2...RH...... - .22475541*T*hobotemplongdate$Ch.2...RH...... - .00683783*hobotemplongdate$Ch.1...Temperature.....C.*hobotemplongdate$Ch.1...Temperature.....C. - .05481717*hobotemplongdate$Ch.2...RH......*hobotemplongdate$Ch.2...RH...... + .00122874*hobotemplongdate$Ch.1...Temperature.....C.*hobotemplongdate$Ch.1...Temperature.....C.*hobotemplongdate$Ch.2...RH...... + .00085282*hobotemplongdate$Ch.1...Temperature.....C.*hobotemplongdate$Ch.2...RH......*hobotemplongdate$Ch.2...RH...... - .00000199*hobotemplongdate$Ch.1...Temperature.....C.*hobotemplongdate$Ch.1...Temperature.....C.*hobotemplongdate$Ch.2...RH......*hobotemplongdate$Ch.2...RH......
```

```{r}
# Heat Index Uncorrected

HIgraph <- ggplot(data=hobotemplongdate, 
    aes(x=longdate, 
    y=HI_uncorrected)) +
    geom_point(size=1, alpha = 1/10)+ theme_bw()+
    facet_grid(cols=vars(location))+ 
    theme(axis.text.x = element_text(angle=65, margin = margin(t=20, r=100)))+
    labs(title="Heat Index Uncorrected", y="Heat Index", x="Date")+
    guides(x = guide_axis(angle=65))+
  scale_x_date(date_breaks = '3 months')
HIgraph

```
```{r}
#HI mean
hobofull <- hobotemplongdate %>%
  tidyr::separate('longdate',
                  into = c('year', 'month', 'day'),
                  sep= '-',
                  remove = FALSE)
head(hobofull)

hobofullHI <- hobofull %>%
  group_by(year,month, day,type,location,longdate)%>%
  summarise(meanHI_uncorrected = mean(HI_uncorrected))

head(hobofullHI)

HImeanplot <- ggplot(hobofullHI, aes(x=longdate, y=meanHI_uncorrected))+
  geom_smooth(aes())+
    geom_point(aes(),alpha = 1/5)+
  facet_grid(cols=vars(location))+
    theme_bw()+
    labs(title= "Daily HI mean", y="Daily mean HI with 95% CI", x="Date")+
    guides(x = guide_axis(angle=65))+
  scale_x_date(date_breaks = '3 months')
   
    
HImeanplot

```

