---
title: "Aug2021 Hobo forest temps"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}

library(ggplot2)
library(tidyverse)

```
## Read in data
Stationary HOBO U23 sensor data from interior and exterior/edge of Lullwater Forest and the Outdoor Activity center - late June - early Aug 2021.
```{r echo=FALSE, include = FALSE}

OACForest <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/OAC1Interior_6-29-21_8-19-21.csv")
head(OACForest)
type<-factor("Forest")
location<-factor("OAC")

#rename variables and add columns for serial no, location, type (forest or edge)
oacF<- data.frame(OACForest,type,location)
head(oacF)

#OAC Edge dataset with type = edge and add location
OACEdge <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/OAC2Exterior_7-7-21_8-19-21.csv")
type<-factor("Edge")
location<-factor("OAC")
oacEdge<- data.frame(OACEdge,type,location)


#Lullwater datasets
LullwaterEdge <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/LullwaterExterior_stonePN4-Emory-4-22-21_0_8-19-21.csv")
type<-factor("Edge")
location<-factor("Lullwater")
lullEdge<- data.frame(LullwaterEdge,type,location)


LullwaterInterior <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/LullwaterInterior1_stoneOH3_6-29-21_8-19-21.csv")
type<-factor("Forest")
location<-factor("Lullwater")
lullF<- data.frame(LullwaterInterior,type,location)

#Fernbank dataset
FernbankInterior <- read.csv("C:/Users/ckeogh/Documents/HOBOware/TempData2021/Aug/FernbankSlope_7.10.21_8-19-21.csv")
type<-factor("Forest")
location<-factor("Fernbank")
fernF<- data.frame(FernbankInterior,type,location)


##Concatenate all 4 datasets by row
alldata <- as.data.frame(rbind(oacF, oacEdge,lullEdge,lullF,fernF))
alldata$longdate<- as.Date(alldata$Date, format = "%Y - %m - %d")
str(alldata)

#https://rpubs.com/cgb-grupstra/moorea-hobo-20190314
#IF need to separate time and date into 2 columns
#hobotemplongdate <- tidyr::separate(hobotemp, 'date',
  #                                  into = c('longdate', 'time'),
  #                                  sep= ' ') 
#head(hobotemplongdate)

```

## Plotting across dates 
using code from https://rpubs.com/cgb-grupstra/moorea-hobo-20190314
```{r echo=FALSE}

ntempgraph <- ggplot(data=alldata, 
    aes(x=longdate, 
    y=TempF, colour=type)) +
    geom_point(size=1, alpha = 1/10)+ theme_bw()+
    facet_grid(type~location)+ 
    theme(axis.text.x = element_text(angle=45, margin = margin(t=20, r=100)))+
    labs(title="Raw temperature data", y="Temperature (°F)", x="Date")+
    guides(x = guide_axis(angle=50))
ntempgraph

```


```{r include=FALSE}

#separating day/month/year 

hobofull <- alldata %>%
  tidyr::separate('longdate',
                  into = c('year', 'month', 'day'),
                  sep= '-',
                  remove = FALSE)
head(hobofull)
```


## Daily temps
Calculating mean temp for each date
```{r mean temps}
hobofullmean <- hobofull %>%
  group_by(year,month, day,type,location,longdate)%>%
  summarise(meantemp = mean(TempF))

head(hobofullmean)

```

## Plotting mean temps
```{r echo=FALSE}

hobofullmean$month <- factor(hobofullmean$month, levels=c("06", "07", "08"))
meanplot <- ggplot(hobofullmean, aes(x=longdate, y=meantemp))+
  geom_smooth(aes(colour=type))+
    geom_point(aes(colour=type),alpha = 1/5)+
  facet_grid(cols=vars(location))+
    theme_bw()+
    labs(title= "Daily temperature means", y="Daily mean temperature (°F) with 95% CI", x="Date")
   
    
meanplot

```

## Calculating daily temp ranges and min/max
```{r temp range}

hobofullrange <- hobofull %>%
  group_by(year, month, day, type, location, longdate)%>%
  summarise(min_temp = min(TempF), max_temp = max(TempF))%>%
  mutate(range = max_temp-min_temp)

head(hobofullrange, digits=5)

```

## Plotting temp ranges
```{r echo=FALSE}

rangeplot <- ggplot(data=hobofullrange, 
                   aes(x=longdate, 
                       y=range, colour=type)) +
  geom_point()+ theme_bw()+
  theme(axis.text.x = element_text(margin = margin(t=5)))+
  facet_grid(cols=vars(location))+
  labs(title="Daily temperature range", y="Daily temperature range (°F) with 95% CI ", x="Date")

rangeplot

```

## Plotting max and min
```{r echo=FALSE}

hobofullrange <- hobofull %>%
  group_by(year, month, day, type, location, longdate)%>%
  summarise(min_temp = min(TempF), max_temp = max(TempF))%>%
 mutate(range = max_temp-min_temp)

minmaxplot <- ggplot(hobofullrange, aes(x=longdate))+
  geom_point(aes(y=min_temp,colour=type))+
    geom_point(aes(y=max_temp,colour=type))+
  facet_grid(cols=vars(location))+
    theme_bw()+
    labs(title= "Daily temperature Min/Max", y="Daily Max and Min temperature (°F) with 95% CI", x="Date")
minmaxplot
   
#    geom_point(aes(colour=type),alpha = 1/5)+
```

## Boxplots - temps per moth by location and type
median, 25%, 75%, and min and max
```{r echo=FALSE}

#hobofullrangemeans<- hobofullrange %>%
#  group_by(year, month,type, location)%>%
#  summarise(meanrange = mean(range))
#nrow(hobofullrangemeans)

hrmplot <- ggplot(hobofullmean, aes(x=month, y=meantemp))+
  geom_boxplot(aes(fill=type)) + theme_bw()+
  facet_grid(cols=vars(location))+
  labs(title= "Monthly average temps", y="Temperature (°C)", x="Date")

hrmplot

```


